Modelling last glacial cycle ice dynamics in the Alps

### 2017-07-13 articles/cycle/alps-cycle.tex
Introduce psi notation for paleo-precipitation.

    Climate forcing driving ice sheet simulations consists of a~present-day
    monthly climatology, $\{T_{\mathrm{m}0}, P_{\mathrm{m}0}\}$, where
    temperatures are modified by offset time series, ${\Delta}T_{\text{TS}}$,
    and lapse-rate corrections, ${\Delta}T_{\text{LR}}$:
    %
    \begin{align}
      T_{\mathrm{m}}(t, x, y) = T_{\mathrm{m}0}(x, y) +
                                {\Delta}T_{\text{TS}}(t) +
                                {\Delta}T_{\text{LR}}(t, x, y) \,,
    \end{align}
    %
    The present-day monthly climatology was bilinearly interpolated from
    near-surface air temperature and precipitation rate fields from
    \citep[WorldClim;][]{Hijmans.etal.2005}, representative of the period 1960
    to 1990. Modern climate of the European Alps is characterised by a
    latitudinal gradient in summer air temperatures, and a longitudinal
    gradient in winter precipitation (Fig.~\ref{fig:inputs}). WorldClim data
    were selected as an input to the ice sheet model because they incorporate
    observations from the dense weather station network of central Europe.
    Besides, last glacial cycle alpine glaciers did not extend over marine
    areas where WorldClim data is missing. Finally, WorldClim data were
    previously used as climate forcing for PISM to model the LGM extent of the
    former Cordilleran ice sheet in good agreement with geological evidence
    along the southern margin \citep{Seguinot.etal.2014} were weather station
    density is lower than in the Alps.

@@

    Finally, in some simulations, precipitation was reduced with temperature in
    order to simulate the potential rarification of atmospheric moisture in
    colder climates. This was done using an empirical relationship derived from
    observed accumulation rates and oxygen isotopes concentrations in the GRIP
    ice core \citep{Dahl-Jensen.etal.1993},
    %
    \begin{align}
      P_{\mathrm{m}}(t, x, y) = P_{\mathrm{m}0}(x, y) \cdot
                                \exp[\zeta{\Delta}T_{\text{LR}}(t, x, y)] \,.
    \end{align}
    %
    with $\zeta=0.169/2.4=0.0704$ \citep{Huybrechts.2002}. This simple
    relationship likely does not reflect the complexity of atmospheric
    circulation changes that governed moisture availability over the Alps
    during the last glacial cycle. Thus other simulations use constant
    precipitation, corresponding to $\zeta=0$. In the rest of this paper, we
    refer to different model runs by the name of the proxy record used for the
    palaeo-temperature forcing. Model runs using paleo-precipitation
    corrections are identified by appending ${\Delta}P$ to their name.

@@

        $\zeta$ & Precipitation factor
                & 0.0704
                & --
                & \citet{Huybrechts.2002} \\

>>>>

    Climate forcing driving ice sheet simulations consists of a~present-day
    monthly climatology, $\{T_{\mathrm{m}0}, P_{\mathrm{m}0}\}$, modified by
    temperature lapse-rate corrections, ${\Delta}T_{\text{LR}}$, temperature
    offset time series, ${\Delta}T_{\text{TS}}$, and time-dependent
    palaeo-precipitation corrections, $\Psi_{\text{PP}}$:
    %
    \begin{align}
      &T_{\mathrm{m}}(t, x, y) = T_{\mathrm{m}0}(x, y) +
                                 {\Delta}T_{\text{LR}}(t, x, y) +
                                 {\Delta}T_{\text{TS}}(t) \,, \\
      &P_{\mathrm{m}}(t, x, y) = P_{\mathrm{m}0}(x, y) \cdot
                                 {\Psi}_{\text{PP}}(t) \,, \\
    \end{align}
    %
    The present-day monthly climatology was bilinearly interpolated from
    near-surface air temperature and precipitation rate fields from
    \citep[WorldClim;][]{Hijmans.etal.2005}, representative of the period 1960
    to 1990. Modern climate of the European Alps is characterised by a
    latitudinal gradient in summer air temperatures, and a longitudinal
    gradient in winter precipitation (Fig.~\ref{fig:inputs}). WorldClim data
    were selected as an input to the ice sheet model because they incorporate
    observations from the dense weather station network of central Europe.
    Besides, last glacial cycle alpine glaciers did not extend over marine
    areas where WorldClim data is missing. Finally, WorldClim data were
    previously used as climate forcing for PISM to model the LGM extent of the
    former Cordilleran ice sheet in good agreement with geological evidence
    along the southern margin \citep{Seguinot.etal.2014} were weather station
    density is lower than in the Alps.

@@

    Finally, in some simulations, precipitation was reduced with temperature in
    order to simulate the potential rarification of atmospheric moisture in
    colder climates. This was done using an empirical relationship derived from
    observed accumulation rates and oxygen isotopes concentrations in the GRIP
    ice core \citep{Dahl-Jensen.etal.1993},
    %
    \begin{align}
      {\Psi}_{\text{PP}}(t) = \exp[\psi{\Delta}T_{\text{TS}}(t)] \,.
    \end{align}
    %
    with $\psi=0.169/2.4=0.0704$ \citep{Huybrechts.2002}. This simple
    relationship likely does not reflect the complexity of atmospheric
    circulation changes that governed moisture availability over the Alps
    during the last glacial cycle. Thus other simulations use constant
    precipitation, corresponding to $\psi=0$. In the rest of this paper, we
    refer to different model runs by the name of the proxy record used for the
    palaeo-temperature forcing. Model runs using paleo-precipitation
    corrections are labelled PP.

@@

        $\psi$  & Precipitation factor
                & 0.0704
                & --
                & \citet{Huybrechts.2002} \\
